\documentclass[final]{nuthesis}
\usepackage{amssymb, amsthm, amsmath, amsfonts}
\usepackage{mathrsfs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{lineno}
\usepackage{listings}
\usepackage{float}
\usepackage{multicol}
\usepackage{xr}

\externaldocument{An_IPM_For_Indeterminate_Growth/An_IPM_For_Indeterminate_Growth}
\externaldocument{Spectral_Properties/Spectral_Properties}
\externaldocument{Estimating_the_Spectral_Radius/Estimating_the_Spectral_Radius}
\externaldocument{Simulations_of_a_Density-Dependent_Growth_Model/Simulations_of_a_Density-Dependent_Growth_Model}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{example}{Example}[section]

\numberwithin{theorem}{section}
\numberwithin{lemma}{section}
\numberwithin{corollary}{section}
\numberwithin{definition}{section}
\numberwithin{equation}{section}

\newcommand{\ms}{\mathscr}
\DeclareMathOperator{\SPAN}{span}
\DeclareMathOperator{\ESSSUP}{ess \, sup}
\DeclareMathOperator{\eig}{eig}

\newcommand{\R}{\mathbb{R}}
\newcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\dd}{\delta}
\newcommand{\ee}{{{\varepsilon}}}

%%% The following commands make the "chi" character look better with subscripts

\makeatletter
\@ifdefinable\@latex@chi{\let\@latex@chi\chi}
\renewcommand*\chi{{\@latex@chi\smash[t]{\mathstrut}}}
\makeatletter

\begin{document}

\frontmatter

\title{Spectral Properties of a Non-Compact Operator in Ecology}
\author{Matt Reichenbach}
\adviser{Professors Richard Rebarber and Brigitte Tenhumberg}
\adviserAbstract{Richard Rebarber and Brigitte Tenhumberg}
\major{Mathematics}
\degreemonth{December}
\degreeyear{2020}
%%
%% For most people the defaults will be correct, so they are commented
%% out. To manually set these, just uncomment and make the needed
%% changes.
%% \college{Your college}
%% \city{Your City}
%%
%% For most people the following can be changed with a class
%% option. To manually set these, just uncomment the following and
%% make the needed changes.
%% \doctype{Thesis or Dissertation}
%% \degree{Your degree}
%% \degreeabbreviation{Your degree abbr.}
%%
%% Now that we know everything we need, we can generate the title page
%% itself.
%%
\maketitle
%%
%% You have a maximum of 350 words for your abstract, which includes
%% your title, name, etc.
%%
%% Required
\input{./abstract}

%% Optional
%% \begin{copyrightpage}
%% \end{copyrightpage}

%% Optional
\begin{dedication}
    \begin{center}
    	To my parents, Kris and Paul.
    \end{center}
\end{dedication}

%%%%%%%%%%%%%%%%%%%
% Acknowledgments
%%%%%%%%%%%%%%%%%%%
\begin{acknowledgments}
	There are many people without whom I never would have been able to write this thesis. First, I want to thank my advisors, Richard Rebarber and Brigitte Tenhumberg. When I applied to graduate school, I never expected to study mathematical biology, but I feel incredibly lucky to have had the chance to work with you both. Brigitte: thank you for always reminding me to think about the biology, and Richard: thank you for all the digressions about music during our meetings. I also want to thank my committee members Yu Jin, Bo Deng, and Kevin Pope, for the many interesting discussions in your classes over the years, and in the Math Bio Seminar.
	
	I want to thank my family for their constant love and support. I would not have been able to finish this project without the support of my parents, so thank you for giving me a place to stay during transitions, your emotional support when I felt career, academic, or existential dread, and for simply chatting on the phone every week. Additionally, over these past five and a half years, I have enjoyed getting to know my sisters Heather and Alyssa as adults; I want you to know that your strength and confidence continue to inspire me. But most of all, I want to thank my niece Violet for letting me leave all of my grown-up worries behind and just play.
	
	I give my heartfelt thanks to all of the people I've met here in Lincoln, because I absolutely would not have been successful here without your friendship. In no particular order, some specific memories I have are: Topology Saturdays with Amadeus Martin, Taran Funk, and Aurora Marks; playing chess with JD Nir, Matthew Bachmann, and Michael DeBellevue; getting outside with Dylan McKnight, Austin Eide, Sara Myers, and Suzi Kang; trivia with Jessie Jamieson, Nick Packauskas, Josh Pollitz, Nathan Poppelreiter, David Lieberman, Kaitlin Tademy, Kathryn Van Etten, Alyssa Whittemore (along with others previously named), and our undaunted host Mike Skiff; making some small statement against systemic racism with Ana Wright, Jake Kettinger (the spookiest witch), and Leilani Pai (my AMS cohost); and because I have no neat-and-tidy category for Su Ji Hong, Rachel Funk, Robert Huben, Paula Egging, and Mohsen Gheibi, thank you for being great friends who were down to just hang out. Of course, this list only scratches the surfaces of my experience here, but they say specificity is the soul of narrative, don't they?
	
	Finally, I want to thank my Brothers Penitent, Karl Stokley and Taylor Britt, for bringing levity to this mad year.
\end{acknowledgments}
%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%
% Grant Information
%%%%%%%%%%%%%%%%%%%
% \begin{grantinfo}
%     % Add any grant info here
% \end{grantinfo}

%%%%%%%%%%%%%%%%%%%
% ToC
%%%%%%%%%%%%%%%%%%%
\tableofcontents

%%%%%%%%%%%%%%%%%%%
% List of Figures
%%%%%%%%%%%%%%%%%%%
\listoffigures
\listoftables

%%%%%%%%%%%%%%%%%%%
% Start of the document
%%%%%%%%%%%%%%%%%%%
\mainmatter

%% Input each chapter as its own file here
\input{An_IPM_For_Indeterminate_Growth/An_IPM_For_Indeterminate_Growth}
\input{Spectral_Properties/Spectral_Properties}
\input{Estimating_the_Spectral_Radius/Estimating_the_Spectral_Radius}
\input{Simulations_of_a_Density-Dependent_Growth_Model/Simulations_of_a_Density-Dependent_Growth_Model}

%% backmatter is needed at the end of the main body of your thesis to
%% set up page numbering correctly for the remainder of the thesis
\backmatter

%% Start the correct formatting for the appendices
%\appendix
%%% Input each appendix here
%\input{./appendix_a}

% \bibliographystyle{alpha}  % or use  abbrv to abbreviate first names and use numerical indices
\bibliographystyle{abbrv}  % or use  abbrv to abbreviate first names and use numerical indices
%% Add your BibTex file here (don't include the .bib)
\bibliography{references}

\end{document}

\endinput
